Computable Poincar\'e--Friedrichs constants\\ for the $L^{p}$ de Rham complex over domains\\ with shellable triangulations

### 2024-08-28 draft.tex
Update on Overleaf.

>>>>

        \item 
        % \todo{Combine all this to get the final transformation.}
        The last simplicial complex that we introduce is called $\calK$,
        and it is the simplicial complex obtained from $\patch_{\calT}(S)$ via barycentric refinement of $S$.

        We let $\calK$ the simplicial complex obtained from $\patch_{\calT}(S)$ via barycentric refinement of $S$. 
        Using our initial observations, we obtain homeomorphisms 
            \Psi_{\calK } : |\calK| \rightarrow \calB,
            \qquad 
            \Psi_{\calR'} : |\calR'| \rightarrow \calB.
        Both $\Psi_{\calK}$ and $\Psi_{\calR'}$ map $T$ onto the same subset of the unit ball.
        Consequently, the image under $\Psi_{\calK}$ of $\overline{|\calK| \setminus T}$
        onto the image under $\Psi_{\calR'}$ of $|\calR^{c}|$.
        It follows that 
            \Xi_{S} := \Psi_{\calK}^{-1} \Psi_{\calR'} \Theta
        is a mapping from $T$ onto $\overline{\patch_{\calT}(S) \setminus T}$.
        We estimate the Jacobians: 
            \| \Jacobian \Xi_{S} \| 
            \sup_{ K \in \calK  }
            \left( 1 + n \kappa(A_{K})^{2} \right)^{\frac 1 2}
            \sigma_{\max}(A_{K}) 
            %
            \sup_{ K \in \calR' } 
            \sqrt{2n} 
            \sigma_{\max}(A_{K}^{-1}) 
            %
            \sup_{ K \in \calR  } \kappa(A_{K})
            \| \Jacobian \Xi_{S}^{-1} \| 
            \sup_{ K \in \calK  }
            \sqrt{2n} 
            \sigma_{\max}(A_{K}^{-1}) 
            %
            \sup_{ K \in \calR' } 
            \left( 1 + n \kappa(A_{K})^{2} \right)^{\frac 1 2}
            \sigma_{\max}(A_{K}) 
            %
            \sup_{ K \in \calR  } \kappa(A_{K})
        Each bisection step reduces the diameter by at most a factor of one half,
        and it reduces the volume always by a factor of one half. 
        Via the estimates in Lemma~\ref{lemma:affinetransform}, we thus find 
            \| \Jacobian \Xi_{S} \| 
            2 \left( \sqrt{n} + 1 \right) 
            \sup_{ K \in \calK  }
            \Ceins{n} \diam(K) 
            \sup_{ K \in \calR' } 
            \Czwei{n} \kappa(K) \diam(K)^{-1}
            \sup_{ K \in \calR  } \kappa(A_{K})
            32 \left( \sqrt{n} + 1 \right) 
            \theta(\calT)
            \Ceins{n} 
            \Czwei{n} 
            \kappa(\calT) 
            \alpha(\calT)
            \| \Jacobian \Xi_{S}^{-1} \| 
            2 \left( \sqrt{n} + 1 \right) 
            \sup_{ K \in \calK  }
            \Czwei{n} \kappa(K) \diam(K)^{-1}
            \sup_{ K \in \calR' } 
            \Ceins{n} \diam(K)
            \sup_{ K \in \calR  } \kappa(A_{K})
            32 \left( \sqrt{n} + 1 \right) 
            \theta(\calT)
            \Ceins{n} 
            \Czwei{n} 
            \kappa(\calT) 
            \alpha(\calT)
        Together with the definition of $\theta(\calT)$, the desired result follows. 
%         \color{red}
%         Each bisection halves the volume of the triangles. Thus,
%         \begin{align*}
%             \| \Jacobian \Xi_{S} \| 
%             2 
%             \sup_{ K \in \patch_{\calT}(S) }
%             2 \sigma_{\max}(A_{K}) 
%             4 
%             \left( \sqrt{n} + 1 \right) \sigma_{\max}(A_{T}^{-1}) 
%             4 
%             \kappa(A_{T})
%             \\&\leq 
%             64
%             \left( \sqrt{n} + 1 \right) 
%             \kappa(A_{T})
%             \sigma_{\max}(A_{T}^{-1}) 
%             \sup_{ K \in \patch_{\calT}(S) }
%             \sigma_{\max}(A_{K}     ) 
%             \| \Jacobian \Xi_{S}^{-1} \| 
%             2 \left( \sqrt{n} + 1 \right) 
%             \sup_{ K \in \patch_{\calT}(S) }
%             \sigma_{\max}(A_{K}) 
%             \sigma_{\max}(A_{T}^{-1}) 
%             4 \kappa(A_{T})
%             \\&\leq 
%             64
%             \left( \sqrt{n} + 1 \right) 
%             \kappa(A_{T})
%             \sigma_{\max}(A_{T}^{-1}) 
%             \sup_{ K \in \patch_{\calT}(S) }
%             \sigma_{\max}(A_{K}     ) 
%             .
%         \end{align*}
%         We use 
%         \begin{gather*}
%             \sup_{ K \in \patch_{\calT}(S) }
%             \sigma_{\max}(A_{K}) 
%             \leq 
%             \Ceins{n} 
%             \sup_{ K \in \patch_{\calT}(S) }
%             \diam(K)
%             \leq 
%             \Ceins{n} 
%             \theta(\calT)
%             \sigma_{\max}(A_{T}) 
%             .
%         \end{gather*}
%         \color{black}
        % That establishes the final estimate. 
%         Suppose that $S = \{v_0\}$ is a single vertex. 
%         We construct a transformation from the vertex star around $S$ onto the Euclidean unit ball. 
%         Without loss of generality, $S$ is at the origin. 
%         For any $T \in \patch_{\calT}(S)$, we let $F_T$ be the transformation. 
%         We define $F$ over $|\patch_{\calT}(S)|$ piecewise. 
%         \todo{Complete}
%

%         \item 
%         Let $\calU$ be the simplicial complex of dimension $n-1$ that contains precisely those facets of $T$ that do not contain $S$.
%         We write $Y := F( |\calU| )$. 
%         \todo{Shape of the outer patch $Y$, star-shaped geometry}
%         
%         \item 
%         We let $\sphere{n}_{\star} := \sphere{n} \setminus \{e_{n+1}\} \subseteq \bbR^{n+1}$ be the unit sphere without the North pole. 
%         Suppose that $\bar x \in \sphere{n}$ is a point on the unit sphere and $\bar x \neq e_{n+1}$. 
%         We define its stereographic projection as 
%         \begin{align*}
%             \pi( \bar x ) 
%             = 
%             \sum_{i=1}^{n} \frac{x_i}{1 - x_{n+1}} e_{i}
%             .
%         \end{align*}
%         This is a mapping $\pi : \sphere{n}_{\star} \rightarrow \bbR^{n}$. 
%         Its inverse is defined as follows: given $y \in \bbR^{n}$, we have 
%         \begin{align*}
%             \pi^{-1}( y )
%             = 
%             \sum_{i=1}^{n} \frac{2 y_i}{1 + \| y \|^2} e_{i}
%             +
%             \frac{1 - \| y \|^2}{1 + \| y \|^2} e_{n+1}
%             .
%         \end{align*}
%         \item 
%         Mirroring the unit sphere along the equator is expressed in stereographic coordinates by the involution 
%         \begin{align*}
%             \mu : \bbR^{n} \rightarrow \bbR^{n},
%             \qquad 
%             y \mapsto \frac{y}{\| y \|}.
%         \end{align*}
%         \item 
%         \begin{align*}
%             G(y) := \| y \| \cdot \phi\left( \frac{y}{\|y\|} \right)^{-1} \frac{y}{\|y\|}
%         \end{align*}
%         \item 
%         We combine these mappings:
%         \begin{align*}
%             Q : \sphere{n} \rightarrow \sphere{n}, \qquad x \mapsto \pi^{-1} G^{-1} \mu G \pi(x)
%         \end{align*}
%         We want to analyze its Lipschitz properties. \todo{Stuff}
%         \item
%         We have constructed a bi-Lipschitz mapping from the Euclidean unit sphere into itself.
%         We use the following result~\cite[Theorem~3]{alestalo2018radial}:
%         if $f : \sphere{n} \rightarrow \sphere{n}$ is any $L$-Lipschitz mapping from the Euclidean unit sphere into itself, 
%         then the radial extension $f : \ball{n} \rightarrow \ball{n}$ is an $L$-Lipschitz mapping from the Euclidean unit ball into itself.
%         Consequently, \todo{write out}
%         \item 
%         In total, we thus have 
%         \begin{align*}
%             Q' := F^{-1} \pi^{-1} G^{-1} \mu G \pi F
%             .
%         \end{align*}

        \item 
        We begin with the following preparatory observations. 
        Consider any $n$-simplex $T$ with vertices $v_0, v_1, \dots, v_n$, where $v_0$ is the origin. 
        We let $A_{T} \in \bbR^{n \times n}$ be the matrix that maps the $i$-th unit vector to $v_i$.
        That matrix maps the reference simplex $\Delta^{n}$ onto $T$. 
        Then the mapping 
            F_T(x) 
            \frac{ \| A_{T}^{-1} x \|_{1} }{ \|x\|_2 } x
        maps $T$ onto a ``simplex with a curved face'' inside the Euclidean unit ball. Its inverse is
            F_T^{-1}(x) 
            \frac{ \|x\|_2 }{ \| A_{T}^{-1} x \|_{1} } x
        We analyze its Lipschitz properties.
        % 
        In what follows, $y \in \bbR^{n}$. 
        we compute the Jacobian at any $x \in \mathring T$:
            \Jacobian F_T(x) \cdot y
            \frac{ \|x\| y - x \|x\|^{-1} x^t y }{\|x\|^2} 
            \left( \vecone \cdot A_{T}^{-1} x \right)
            \frac{x}{\|x\|} \left( \vecone \cdot A_{T}^{-1} y \right)
            \\&= 
            \left( \Id - \|x\|^{-2} x \otimes x^t \right)
            \frac{ 1 }{ \|x\| } 
            \left( \vecone \cdot A_{T}^{-1} x \right)
            \frac{x}{\|x\|} \left( \vecone \cdot A_{T}^{-1} x \right)
            \\&= 
            \left( A_{T}^{-t} \vecone \cdot \hat x \right)
            \left( \Id - \hat x \otimes \hat x^t \right) y
            \left( A_{T}^{-t} \vecone \cdot y \right) \hat x 
        This is an orthogonal sum of two vectors. Its norm satisfies:
            \| \Jacobian F_T(x) \cdot y \|^{2}
            \left| A_{T}^{-t} \vecone \cdot \hat x \right|^{2}
            \| \left( \Id - \hat x \otimes \hat x^t \right) y \|^{2}
            \left| A_{T}^{-t} \vecone \cdot y \right|^{2} \| \hat x \|^{2}
        The Euclidean operator norm satisfies 
            \| \Jacobian F_T(x) \| 
            \sqrt 2 \left\| A_{T}^{-t} \vecone \right\|_{2}
            \sqrt{2n} \cdot \sigma_{\max}(A_{T}^{-1}) 
        On the other hand, we can rewrite the Jacobian as:
            \Jacobian F_T(x) \cdot y
            \left( A_{T}^{-t} \vecone \cdot \hat x \right)
            y
            \left( A_{T}^{-t} \vecone \cdot \hat x \right)
            \left( \hat x \otimes \hat x^t \right) y
            \left( A_{T}^{-t} \vecone \cdot y \right) \hat x 
            \\&= 
            \left( A_{T}^{-t} \vecone \cdot \hat x \right)
            y
            \hat x 
                \left( A_{T}^{-t} \vecone \cdot \hat x \right) \hat x^{t}
                \vecone^{t} A_{T}^{-1} 
            \right) y
            \\&= 
            \left( A_{T}^{-t} \vecone \cdot \hat x \right)
            y
            \hat x 
                \left( A_{T}^{-t} \vecone \cdot \hat x \right) \hat x
                A_{T}^{-t} \vecone
            \right)^{t} y
        Its inverse is found via the Woodbury formula: 
        % Its inverse is easily verified to be:
            \Jacobian F_T(x)^{-1}
            \left( A_{T}^{-t} \vecone \cdot \hat x \right)^{-1}
            \Id
            \left( A_{T}^{-t} \vecone \cdot \hat x \right)^{-2}
            \hat x 
                \left( A_{T}^{-t} \vecone \cdot \hat x \right) \hat x
                A_{T}^{-t} \vecone
            \right)^{t} 
            \\&= 
            \left( A_{T}^{-t} \vecone \cdot \hat x \right)^{-1}
                \Id
                \hat x 
                    % \left( A_{T}^{-t} \vecone \cdot \hat x \right)^{-1}
                    % \left( A_{T}^{-t} \vecone \cdot \hat x \right) 
                    \hat x
                    \left( A_{T}^{-t} \vecone \cdot \hat x \right)^{-1}
                    A_{T}^{-t} \vecone
                \right)^{t}
            \\&= 
            \left( \vecone \cdot A_{T}^{-1} \hat x \right)^{-1}
                \Id
                \hat x 
                    % \left( A_{T}^{-t} \vecone \cdot \hat x \right)^{-1}
                    % \left( A_{T}^{-t} \vecone \cdot \hat x \right) 
                    \hat x
                    \left( \vecone \cdot A_{T}^{-1} \hat x \right)^{-1}
                    A_{T}^{-t} \vecone
                \right)^{t}
            \\&= 
            \left( \vecone \cdot A_{T}^{-1} \hat x \right)^{-1}
                \Id - \hat x \otimes \hat x^{t}
                \left( \vecone \cdot A_{T}^{-1} \hat x \right)^{-1}
                \hat x \otimes \vecone^{t} A_{T}^{-1}
        The inverse acts via multiplication with $\left( \vecone \cdot A_{T}^{-1} \hat x \right)^{-1}$ on the span of $\hat x$.
        We notice the upper bound 
            \left( \vecone \cdot A_{T}^{-1} \hat x \right)^{-1}
            \| A_{T}^{-1} \hat x \|^{-1}_{1}
            \| A_{T}^{-1} \hat x \|^{-1}_{2}
            \sigma_{\min}( A_{T}^{-1} )^{-1}
            \sigma_{\max}( A_{T} )
        This holds because $A_{T}^{-1} \hat x$ is in the positive quadrant.
        %, and its scalar product with $\vecone$ is minimized when it is colinear with a coordinate axis.
        If $y \in \bbR^{n}$ is orthogonal to $\hat x$, then 
            \| \Jacobian F_T(x)^{-1} \cdot y \|^{2}
%             \left( \vecone \cdot A_{T}^{-1} \hat x \right)^{-2}
%             \left\| 
%                 y
%             \right\|^{2}
%             \left( \vecone \cdot A_{T}^{-1} \hat x \right)^{-2}
%             \left| 
%                 \left( \vecone \cdot A_{T}^{-1} \hat x \right)^{-1}
%                 \vecone^{t} A_{T}^{-1} y 
%             \right|^{2}
%             \\&\leq 
            \left( \vecone \cdot A_{T}^{-1} \hat x \right)^{-2}
                y
            \right\|^{2}
            \frac{ 
                \left( \vecone \cdot A_{T}^{-1} \hat x \right)^{-1}
                \vecone^{t} A_{T}^{-1} y 
            \right|^{2}
            }{
            \left( \vecone \cdot A_{T}^{-1} \hat x \right)^{2}
            }
        The operator norm of that inverse is:
            \| \Jacobian F_T(x)^{-1} \|
            \sigma_{\max}( A_{T} )
            \left( 1 + n \frac{ \sigma_{\max}(A_{T})^{2} }{ \sigma_{\min}(A_{T})^{2} } \right)^{\frac 1 2}
        Since the domain and codomain of $F_{T}$ are convex, 
        this also implies bounds on the Lipschitz constants of $F_{T}$ and its inverse.
